---
title: "Course Project pt2"
author: "Zach Dungan"
date: "3/3/2022"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 2: Inferential Data Analysis

Load libraries and ToothGrowth data

``` {r}
library(ggplot2)
library(datasets)

data("ToothGrowth")
```

**View info about ToothGrowth data**

``` {r}
str(ToothGrowth)
```

``` {r}
head(ToothGrowth)
```

Summary statistics
``` {r}
summary(ToothGrowth)
```
Correct 'dose' varibale
``` {r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

**Data Analysis**

``` {r}
supp_mean <- split(ToothGrowth$len, ToothGrowth$supp)
sapply(supp_mean, mean)

supp_dose <- split(ToothGrowth$len, ToothGrowth$dose)
sapply(supp_dose, mean)
```

Graph of effect of supplement on tooth length:
``` {r}
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + xlab("Supplement") + ylab("Length")
```

Graph of effect of supplement on tooth length:
``` {r}
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose)) + xlab("Dose") + ylab("Length")
```

**Statistical Tests**

Create easy to use variables
``` {r}
supp <- ToothGrowth$supp
dose <- ToothGrowth$dose
len <- ToothGrowth$len
```


T-Test #1

``` {r}
# Check for differences in supplement type
t.test(len[supp == "VC"], len[supp == "OJ"])
```

The p-value is about 0.06, meaning we cannot reject the null hypothesis that different supplements don't have an effect on tooth length.

T-Test #2
``` {r}
# Check for differences in dosage type
# In this test, use 0.5 and 1.0
t.test(len[dose == 0.5], len[dose == 1.0])
```

The p-value is very close to zero, which shows a considerable impact on tooth length, meaning we can reject the null hypothesis. 

For thoroughness, the tests between [0.5 and 2.0] and [1.0 and 2.0] are below, with similar results

``` {r}
# 0.5 and 2.0
t.test(len[dose == 0.5], len[dose == 2.0])

# 1.0 and 2.0
t.test(len[dose == 1.0], len[dose == 2.0])
```

**Conclusions**

1. The supplement type, statistically speaking, has no significant effect on tooth length.
2. The dosage amount has a significant effect on tooth length


















